ABM simulating short-term mating behaviors

### 2022-05-13 submission/manuscript.tex
minor language edits of the manuscript, tracked the model directory, close #18

\hypertarget{this-is-an-awesome-title}{%
\section*{\texorpdfstring{\textbf{This is an awesome
title}}{This is an awesome title}}\label{this-is-an-awesome-title}}
\addcontentsline{toc}{section}{\textbf{This is an awesome title}}

@@

Yurun Ying\textsuperscript{1}, Pekka
Santtila\textsuperscript{1,2,\(\dagger\)}

@@

\textsuperscript{2} NYU-ECNU Institute for Social Development at NYU
Shanghai

@@

\textsuperscript{\(\dagger\)} To whom correspondence should be
addressed:

@@

Gender differences in short-term mating behaviors have been a
well-documented phenomenon in human sexuality research. Existing studies
usually conflate gender differences in preferences for short-term mating
with differences in sexual behaviors, which is theoretically dubious.
Using an agent-based model, the present study investigated whether men
and women's different mating preferences resulted in any gender
differences in short-term mating behaviors, and if they did, under what
circumstances. The results suggested that heterosexual men and women had
the same average number of short-term mating experiences and short-term
mates even when gender differences in preferences fort-term mating
existed. Gender differences in mating behaviors only emerged when
heterosexual men and women in the mating pool were considered. Moreover,
gay men had a higher average number of both outcomes than lesbian women
and than heterosexual men under such a circumstance. These results
suggest that theoretically speaking, gender differences in short-term
mating behaviors only occurred among particular populations, or when
men's preferences for short-term mating are not constrained by those of
women. Suggestions for future research in human mating psychology and
behaviors were provided.

@@

Intuitive as this line of reasoning is, there are good reasons to doubt
its soundness. This is because any heterosexual sexual encounter
involves both a man and a woman (the frequency of encounters involving
more than two persons is negligible, (e.g., Herbenick et al., 2017). A
new short-term mating experience for a man is, therefore, also a new
experience for a woman, and the same logic applies to counting a new
short-term mate. As a result, the psychological differences in
short-term mating preferences may not result in behavioral ones in the
heterosexual case (Archer, 2019).

The present study took gender differences in mating preferences as an
assumption, which was proposed by sexual strategy theory (Buss \&
Schmitt, 1993) and repeatedly found support by empirical studies (e.g.,
Schmitt, 2003; Walter et al., 2020). Using an agent-based model, we
investigated whether men and women's different mating preferences
resulted in any gender differences in short-term mating behaviors,
specifically, in the number of short-term mating experiences and
short-term mates, and if they did, under what circumstances.

@@

Gender differences in mating strategies have been studied in the light
of sexual strategy theory (Buss \& Schmitt, 1993). It posits that since
the minimum investment that men devote to their offspring (contribution
of sperm through one sexual act) is lower than that of women (gestation,
labor, and lactation), men tend to have short-term mating as a larger
part of their mating strategy than women do. This is because women's
high minimum investment leads to higher opportunity cost (vs.~men) if
they have sex with a suboptimal partner and give birth to offspring with
a low survival chance. As a result, men may have a greater interest in
short-term mating, desire a larger number of short-term mates in a given
period, and be less selective with respect to accepting a potential mate
(Buss \& Schmitt, 1993).

@@

As for mating standards, men are more likely to accept someone as a
potential short-term mate. For example, studies using U.S. college
samples found that the minimum percentile ranks that men found
acceptable for a potential short-term sexual partner were lower than for
women both overall and on individual traits (e.g., social status,
attractiveness) (Kenrick et al., 1990; Regan, 1998). A study also found
that when presented with identical descriptions of potential mates, men
on average rated them as more desirable than women did (Wiederman \&
Dubois, 1998).

Some evidence shows that sexual strategy theory also predicts mating
preferences among gay men and lesbian women. A survey study using a
community sample from the U.S. found that gay men were more interested
in short-term mating than lesbian women and that this gender difference
was comparable to that existing among heterosexual individuals (Bailey
et al., 1994). An indirect piece of evidence on differential standards
for short-term mates comes from a recent study finding that
significantly more gay men than lesbian women reported to have accepted
a casual sexual offer from a same-gender person (Matsick et al., 2021).
Past studies suggest that the gender difference in the acceptance rate
of casual sexual offers may originate from men and women's differential
standards for short-term mates (Conley et al., 2011; Hald \&
Høgh-Olesen, 2010). Thus, as a postulation, gay men and lesbian women
may also have different standards for short-term mates.

@@

Since heterosexual sex involves both a man and a woman in most cases,
men's interest in short-term mating can be constrained by women's
preference for long-term relationships (Archer, 2019; Symons, 1979).
This is because men's short-term mating preferences can only translate
into behaviors when there are women willing to have sex with them. When
short-term mating occurs, the number of short-term mating encounters and
short-term mates is essentially the same for men and women (although
this does not necessarily mean that the total number of men and women
who have ever had short-term mating must be equal). Therefore, we would
expect to observe no gender difference in short-term mating behaviors
among heterosexual individuals even when there were gender differences
in mating preferences.

As a comparison, in the cases of gay men and lesbian women, men's
preferences are not constrained by women's, but only by those of other
men, who, arguably, have more similar preferences. This would allow for
a more direct behavioral expression of men's mating preferences. The
notion that gay men have less restricted preferences was borne out in
the proportion of individuals who have engaged in extradyadic sex
(Peplau et al., 2004). A study showed that the proportions of
heterosexual men and women who have engaged in extradyadic sex were 26\%
and 21\%, respectively, while the statistics among gay men and lesbian
women were 82\% and 28\%, respectively (Peplau et al., 2004). Therefore,
we would theoretically expect to observe gender differences in
short-term mating behaviors among gay men and lesbians. Moreover, we
would also expect to find that gay men, as compared to heterosexual men,
engaged in more short-term mating behaviors due to the lessened
constraint on their mating preferences.

@@

By using a spatial agent-based model, the present study investigated
whether men and women's different mating preferences resulted in any
gender differences in short-term mating behaviors, and if they did,
under what circumstances. The interest in short-term mating was modeled
as an individual's likelihood of deciding to have short-term mating at
each time step. The standards for short-term mates were modeled as the
minimum desirability of a potential mate with whom an individual was
willing to have sex. Short-term mating behaviors were operationally
defined as the number of short-term mating experiences and the number of
past short-term mates.

We modeled this process among both heterosexual individuals and gay men
and lesbian women to examine whether constraints set by the opposite
sex's preferences would make a difference in short-term mating
behaviors. Individuals' sexual orientation was conceptualized in terms
of behaviors only in our model. Heterosexual men had short-term mating
or formed a long-term relationship with women, while gay men only with
other men, and vice versa for women.

@@

The model measured two outcomes: (1) the number of short-term mating
experiences of men and women; (2) the number of short-term mates of men
and women. Additionally, we had also documented the number of men and
women in the mating pool (i.e., who had ever engaged in short-term
mating); The average number of short-term mating experiences and
short-term mates were calculated by taking the average among the whole
population of men/women (\emph{N\textsubscript{m}} =
\emph{N\textsubscript{w}} = 150) and among those who had engaged in
short-term mating.

See the Supplemental Materials for the overview, design concepts, and
details (ODD) protocol of the model, which includes detailed scheduling
and parameterization. The model can be downloaded from the Github online
repository {[}link masked for peer review{]}. The model was programmed
in Netlogo 6.2.1 (Wilensky, 1999).

@@

Using the agent-based model, the present study conducted two 2 (gender
difference vs.~no difference in short-term mating likelihood) x 2
(gender difference vs.~no difference in mating standard) experiments.
Experiment 1 was run among heterosexual agents who only engaged in
short-term or long-term relationships with agents of the opposite
gender. Experiment 2 was run among gay men and lesbian women who only
engaged in short-term or long-term relationships with agents of the same
gender.

Gender differences were modeled in agents' likelihood of engaging in
short-term mating and the standard for short-term mates. The values of
these parameters were set based on empirical findings in human mating
psychology. We compared the conditions where there were gender
differences in mating preferences with counterfactual conditions where
gender differences were missing. When there was a gender difference in
short-term mating likelihood, men had a 40\% likelihood of deciding to
engage in short-term mating upon meeting a potential mate, while women
had a likelihood of 25\% (Buss \& Schmitt, 1993; Schmitt et al., 2001;
Schmitt, 2003). When there was no gender difference in short-term mating
likelihood, both women and men had a likelihood of 25\%. When there was
a gender difference in mating standards, men had a mating standard of 3
(the minimum mate value of a potential partner with whom an agent is
willing to have short-term mating, highest possible value = 10), while
women had a standard of 5 (Kenrick et al., 1990; Regan, 1998). When
there was no gender difference in mating standards, both men and women
had a standard of 5.

@@

At each time step, men and women moved and decided on short-term mating
as the agents did in the heterosexual procedures. Half of the agents (an
equal number of men and women) were also given the initiator status. The
initiators checked to see whether there were other agents at the same
location. The rest of the procedures were identical to the heterosexual
procedures, except that the agents only chose those with the same gender
as themselves as potential short-term mates or long-term partners.

@@

The model was run for 1,000 time steps in each simulation. We ran 10,000
simulations for each experiment, and 2,500 simulations for each
condition. We controlled for initializing random seeds in the model
runs. All simulations were run using Netlogo 6.2.1 (Wilensky, 1999).

@@

\hypertarget{gender-differences-in-short-term-mating-behaviors-in-each-experiment}{%
\subsection{Gender differences in short-term mating behaviors in each
experiment}\label{gender-differences-in-short-term-mating-behaviors-in-each-experiment}}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=80mm]{figures/fig1_men_vs_women.pdf}
  \caption{\textbf{Short-term mating behaviors of men and women after 1,000 time steps in the model when gender differences existed in mating preferences.} Violin plots summarizing the outcome variables separately for heterosexual individuals and gay men and lesbian women. Plot (A) shows the average number of short-term mating experiences, and plot (B) shows the average number of short-term mates. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar. Statistics were calculated using the full population of men and women in the model (\textit{N\textsubscript{m}} = \textit{N\textsubscript{w}} = 150).}
\end{figure}

@@

In Experiment 1, however, when the means of the outcome variables were
calculated among men and women who were in the mating pool (i.e., those
with a non-zero number of short-term mating experiences) \textbf{(the
number of individuals in the mating pool: M\textsubscript{m},
SD\textsubscript{m}; M\textsubscript{w}, SD\textsubscript{w})}, there
were gender differences in short-term mating behaviors when there were
gender differences in mating preferences. The average number of
short-term mating experiences was higher among heterosexual men
(\emph{M} = 3.78, \emph{SD} = 0.50) than among heterosexual women
(\emph{M} = 2.79, \emph{SD} = 0.35), \emph{t}(4,998) = -81.53, \emph{p}
\textless{} .001, Cohen's \emph{d} = 2.31. Likewise, the average number
of short-term mates was higher among heterosexual men (\emph{M} = 2.12,
\emph{SD} = 0.17) than among heterosexual women (\emph{M} = 1.56,
\emph{SD} = 0.11), \emph{t}(4,998) = -135.85, \emph{p} \textless{} .001,
Cohen's \emph{d} = 3.84.

\hypertarget{comparing-short-term-mating-behaviors-of-heterosexual-and-gay-men}{%
\subsection{Comparing short-term mating behaviors of heterosexual and
gay
men}\label{comparing-short-term-mating-behaviors-of-heterosexual-and-gay-men}}

@@

In an exploratory manner, we also compared across conditions to see
which of the two dimensions of preferences for short-term mating
contributed to gender differences in sexual behaviors.

Among heterosexual individuals, gender differences in short-term mating
behaviors, as calculated among individuals in the mating pool, emerged
when men and women had different standards for short-term mates.
Heterosexual men in the mating pool had a higher average number of
short-term mating experiences and short-term mates than women in the
mating pool, even when no gender difference existed in short-term mating
likelihood (experiences: \emph{M\textsubscript{m}} = 2.82,
\emph{SD\textsubscript{m}} = 0.36, \emph{M\textsubscript{w}} = 2.21,
\emph{SD\textsubscript{w}} = 0.28, Cohen's \emph{d} = 1.89; partners:
\emph{M\textsubscript{m}} = 1.76, \emph{SD\textsubscript{m}} = 0.15;
\emph{M\textsubscript{w}} = 1.38, \emph{SD\textsubscript{w}} = 0.09,
Cohen's \emph{d} = 3.13). In comparison, virtually no gender differences
in short-term mating behaviors existed when men and women had the same
mating standards (Table 1).

@@

Among gay men and lesbian women, gender differences in short-term mating
behaviors emerged either when they had different short-term mating
likelihood or when their mating standards were different. Gay men had a
higher average number of short-term mating experiences and short-term
mates when they had a higher short-term mating likelihood than lesbian
women (experiences: \emph{M\textsubscript{m}} = 0.80,
\emph{SD\textsubscript{m}} = 0.23, \emph{M\textsubscript{w}} = 0.32,
\emph{SD\textsubscript{w}} = 0.10, Cohen's \emph{d} = 2.67; partners:
\emph{M\textsubscript{m}} = 0.52, \emph{SD\textsubscript{m}} = 0.13,
\emph{M\textsubscript{w}} = 0.25, \emph{SD\textsubscript{w}} = 0.07,
Cohen's \emph{d} = 2.60), or when they had a lower mating standards
(experiences: \emph{M\textsubscript{m}} = 1.05,
\emph{SD\textsubscript{m}} = 0.18, \emph{M\textsubscript{w}} = 0.32,
\emph{SD\textsubscript{w}} = 0.10, Cohen's \emph{d} = 5.10; partners:
\emph{M\textsubscript{m}} = 0.82, \emph{SD\textsubscript{m}} = 0.12,
\emph{M\textsubscript{w}} = 0.25, \emph{SD\textsubscript{w}} = 0.07,
Cohen's \emph{d} = 5.78). In comparison, virtually no gender differences
in short-term mating behaviors existed when gay men and lesbian women
had the same short-term mating likelihood and mating standards (Table
2).

@@

The present study aimed at using agent-based modeling to investigate
whether men and women's different mating preferences resulted in any
gender differences in short-term mating behaviors, specifically, in the
number of short-term mating experiences and short-term mates, and if
they did, under what circumstances. We raised two hypotheses: 1) there
would be gender differences in short-term mating behaviors among gay men
and lesbian women, with gay men being engaged in more such behaviors,
and 2) gay men would engage in more short-term mating behaviors as
compared to heterosexual men. The results from 1,000 time steps in a
model simulating men and women's mating behaviors supported our
hypotheses. First of all, as compared to lesbian women, gay men had a
higher average number of short-term mating experiences and short-term
mates. In contrast, we found no gender differences in short-term mating
behaviors between heterosexual men and women. Secondly, gay men also had
a higher average number of short-term mating experiences and short-term
mates as compared to heterosexual men.

As we expected, heterosexual men and women did not differ in short-term
mating behaviors despite their different mating preferences. This was
because the total numbers of short-term mating experiences and
short-term mates of heterosexual men and women were exactly the same.
Since the sex ratio was 1:1 in our model, the average number of
short-term mating experiences and short-term mates must be equal between
men and women as well. However, we did find that among individuals in
the mating pool, men engaged in more short-term mating behaviors as
compared to women. This was because there were more women than men in
the mating pool (i.e., had ever engaged in short-term mating behaviors),
resulting in lower averages among heterosexual women despite equal
numbers of experiences and mates in total.

Moreover, gender differences in short-term mating behaviors emerged when
heterosexual men and women had different standards for short-term mates,
but not when they had different likelihood of engaging in short-term
mating. When women had a higher standard than men did, less men than
women in the population were above a potential partner's standards and
thus had a chance to have sex with them. This contributed to the unequal
number of men and women in the mating pool, which led to gender
differences in short-term mating behaviors.

In the light of these results, the empirical observation of gender
differences in short-term mating behaviors among heterosexual
individuals (e.g., Petersen \& Hyde, 2010) appear to be perplexing
because this is theoretically illogical (e.g., Gurman, 1989). Possible
explanations for these empirical results have to highlight caveats in
the observation process. One possibility is that there was sampling bias
in the observations (e.g., Wiederman \& Dubois, 1998) since surveys
regarding short-term mating behaviors may tend to attract individuals
who already engage in such behaviors. Another possibility is that
heterosexual men and women tend to report short-term mating behaviors
differently. This difference may be a result of social desirability bias
(Alexander \& Fisher, 2003) or due to men and women's different
estimation strategies (e.g., men tend to round up) (Brown \& Sinclair,
1999).

Among gay men and lesbian women, large gender differences in short-term
mating behaviors existed when men and women had different mating
preferences, which was in line with empirical observations (e.g., Peplau
et al., 1997, 2004). This was probably because the number of short-term
mating experiences and short-term mates no longer counted towards men
and women simultaneously. Any gender differences in mating preferences
would result in differences in behaviors. A closer look at the results
did support this postulation. Either a difference in the likelihood of
short-term mating or a difference in mating standards alone could
contribute to gender differences in mating behaviors. This was perhaps
because the former increased the probability of both parties of a given
gay couple deciding to have short-term mating, and the latter increased
the probability of both of them meeting each other's standards, as
compared to the case of a given lesbian couple.

Supporting the idea that gay men's preferences for short-term mating
were not constrained by those of women, we found that gay men engaged in
more short-term mating behaviors as compared to heterosexual men. This
was consistent with the empirical literature (e.g., Peplau et al.,
1997). Interestingly, gay men and heterosexual men had the same
likelihood of short-term mating and the same standards for short-term
mates in our model. The only difference was a change in the preferences
in their potential partners. When men's partners had a stronger
preferences for short-term mating (i.e., having men vs.~having women as
potential partners), men also appeared to engage in more short-term
mating behaviors.

>>>>

\hypertarget{an-agent-based-model-of-gender-differences-in-short-term-mating-behaviors-as-a-result-of-mating-preferences}{%
\section*{An Agent-based Model of Gender Differences in Short-term
Mating Behaviors as a Result of Mating
Preferences}\label{an-agent-based-model-of-gender-differences-in-short-term-mating-behaviors-as-a-result-of-mating-preferences}}
\addcontentsline{toc}{section}{An Agent-based Model of Gender
Differences in Short-term Mating Behaviors as a Result of Mating
Preferences}

@@

Yurun Ying\textsuperscript{1}, Jan Antfolk\textsuperscript{2}, Pekka
Santtila\textsuperscript{1,3,\(\dagger\)}

@@

\textsuperscript{2} xxxx

\textsuperscript{3} NYU-ECNU Institute for Social Development at NYU
Shanghai

@@

\textsuperscript{\(\dagger\)} To whom correspondence should be
addressed: \url{pekka.santtila@nyu.edu}

@@

Gender differences in short-term mating behaviors have been a
well-documented phenomenon in human sexuality research. Existing studies
usually conflate gender differences in mating preferences with
differences in sexual behaviors, which is theoretically dubious. Using
an agent-based model, the present study investigated the circumstances
under which men and women's differential preferences for short-term
mating would result in gender differences in short-term mating
behaviors. The results from the model suggested that heterosexual men
and women had the same average number of short-term mating experiences
and short-term mates even when men had stronger preferences for
short-term mating. Gender differences in mating behaviors only emerged
when heterosexual men and women in the mating pool (i.e., those with a
non-zero number of short-term mating experiences) were considered.
Moreover, gay men had a higher average number of both outcomes than
lesbian women as well as than heterosexual men when men had stronger
preferences for short-term mating. These results suggest that
theoretically speaking, even when gender differences in mating
preferences existed, those in short-term mating behaviors only occurred
among particular populations, or when men's preferences for short-term
mating are not constrained by those of women. Suggestions for future
research in human mating psychology and behaviors were provided.

@@

Intuitive as this line of reasoning is, there are good reasons to doubt
its soundness. This is because any heterosexual sexual encounter
involves both a man and a woman (the frequency of encounters involving
more than two persons is negligible, e.g., Herbenick et al., 2017). A
new short-term mating experience for a man is, therefore, also a new
experience for a woman, and the same logic applies to counting a new
short-term mate. As a result, the psychological differences in
short-term mating preferences may not result in behavioral ones in the
heterosexual case (Archer, 2019).

The present study took gender differences in mating preferences as a
presumption, which was conceptualized by sexual strategy theory (Buss \&
Schmitt, 1993) and repeatedly found support by empirical studies (e.g.,
Schmitt, 2003; Walter et al., 2020). Using an agent-based model, we
investigated whether men and women's differential preferences for
short-term mating resulted in any gender differences in short-term
mating behaviors, specifically, in the number of short-term mating
experiences and short-term mates, and if they did, under what
circumstances.

@@

Gender differences in mating preferences have been studied in the light
of sexual strategy theory (Buss \& Schmitt, 1993). It posits that since
the minimum investment that men devote to their offspring (contribution
of sperm through one sexual act) is lower than that of women (gestation,
labor, and lactation), men tend to have short-term mating as a larger
part of their mating strategy than women do. This is because women's
high minimum investment leads to higher opportunity cost (vs.~men) if
they have sex with a suboptimal partner and give birth to offspring with
a low survival chance. As a result, men may have a greater interest in
short-term mating, desire a larger number of short-term mates in a given
period, and be less selective with respect to accepting a potential mate
(Buss \& Schmitt, 1993).

@@

As for mating standards, men are less selective in accepting someone as
a potential short-term mate. For example, studies using U.S. college
samples found that the minimum percentile ranks that men found
acceptable for a potential short-term sexual partner were lower than
those accepted by women, both in terms of overall desirability and
individual traits (e.g., social status, attractiveness) (Kenrick et al.,
1990; Regan, 1998). A study also found that when presented with
identical descriptions of potential mates, men on average rated them as
more desirable than women did (Wiederman \& Dubois, 1998).

Some evidence shows that gender differences in mating preferences also
existed between gay men and lesbian women. A survey study using a
community sample from the U.S. found that gay men were more interested
in short-term mating than lesbian women were, and that this gender
difference was comparable to that existing among heterosexual
individuals (Bailey et al., 1994). An indirect piece of evidence on
differential standards for short-term mates comes from a recent study
finding that significantly more gay men than lesbian women reported to
have accepted a casual sexual offer from a same-gender person (Matsick
et al., 2021). Past studies suggest that the gender difference in the
acceptance rate of casual sexual offers may originate from men and
women's differential standards for short-term mates (Conley et al.,
2011; Hald \& Høgh-Olesen, 2010). Thus, as a postulation, gay men may
also have lower standards than lesbian women for short-term mates.

@@

Since heterosexual sex involves both a man and a woman in most cases,
men's interest in short-term mating can be constrained by women's
preference for long-term relationships (Archer, 2019; Symons, 1979).
This is because men's short-term mating preferences can only translate
into behaviors when there are women willing to have sex with them. When
a new short-term mating encounter occurs, it counts towards men's total
number of short-term mating encounters as well as towards women's.
Similarly, a new short-term mate counts towards both men and women's
total number of short-term mates. Therefore, the total number of
short-term mating experiences and short-term mates must be equal between
heterosexual men and women on a population level. However, it is
important to note that although heterosexual men's total number of
short-term mating is equal to that of women, this does not necessarily
mean that the total number of men and women who have ever had short-term
mating must be equal. It is possible that there is a smaller proportion
of men (vs.~the proportion of women) who contribute to the total number
of short-term mating. Following this line of reasoning, we would expect
to observe no gender difference in short-term mating behaviors among
heterosexual individuals even when there were gender differences in
mating preferences.

As a comparison, in the cases of gay men and lesbian women, men's
preferences are not constrained by women's, but only by those of other
men, who, arguably, have more similar preferences. This would allow for
a more direct behavioral expression of men's mating preferences. The
notion that gay men have less restricted preferences was borne out in
the proportion of individuals who have engaged in extradyadic sex
(Peplau et al., 2004). A study showed that the proportions of
heterosexual men and women who have engaged in extradyadic sex were 26\%
and 21\%, respectively, while the statistics among gay men and lesbian
women were 82\% and 28\%, respectively (Peplau et al., 2004). Therefore,
we would theoretically expect to observe gender differences in
short-term mating behaviors between gay men and lesbian women. Moreover,
we would also expect to find that gay men, as compared to heterosexual
men, engaged in more short-term mating behaviors due to the lessened
constraint on their mating preferences.

@@

By using a spatial agent-based model, the present study investigated
whether men and women's different preferences for short-term mating
resulted in any gender differences in short-term mating behaviors, and
if they did, under what circumstances. Preferences for short-term mating
were operationally defined as the interest in short-term mating and
standards for short-term mates (Buss \& Schmitt, 1993). Short-term
mating behaviors were operationally defined as the number of short-term
mating experiences and the number of past short-term mates (Petersen \&
Hyde, 2010).

We modeled this process among both heterosexual individuals and gay men
and lesbian women to examine whether constraints set by the opposite
sex's preferences would affect short-term mating behaviors. Individuals'
sexual orientation was conceptualized in terms of behaviors only in our
model. Heterosexual men had short-term mating or formed a long-term
relationship only with women, while gay men only with other men, and
vice versa for women.

@@

The model measured two outcomes: (1) the number of short-term mating
experiences of men and women; (2) the number of short-term mates of men
and women. Additionally, we also measured the number of men and women in
the mating pool (i.e., those with a non-zero number of short-term mating
experiences). The average numbers of short-term mating experiences and
short-term mates were calculated by taking the average among the whole
population of men/women (\emph{N\textsubscript{m}} =
\emph{N\textsubscript{w}} = 150) and among those who had engaged in
short-term mating.

See the Supplemental Materials for the overview, design concepts, and
details (ODD) protocol of the model, which includes detailed scheduling
and parameterization. The model can be downloaded from the Open Science
Framework {[}link masked for peer review{]}. The model was programmed in
Netlogo 6.2.1 (Wilensky, 1999).

@@

Using the agent-based model, the present study conducted two 2 (gender
difference vs.~no difference in the interest in short-term mating
likelihood) x 2 (gender difference vs.~no difference in mating
standards) experiments. Experiment 1 was run among heterosexual agents
who only engaged in short-term or long-term relationships with agents of
the opposite gender. Experiment 2 was run among gay men and lesbian
women who only engaged in short-term or long-term relationships with
agents of the same gender.

The interest in short-term mating was modeled as an agent's likelihood
of deciding to have short-term mating at each time step. The standards
for short-term mates were modeled as the minimum desirability of a
potential mate with whom an agent was willing to have sex. The values of
the two parameters were set based on empirical findings in human mating
psychology. When there was a gender difference in short-term mating
likelihood, men had a 40\% likelihood of deciding to engage in
short-term mating, while women had a likelihood of 25\% (Buss \&
Schmitt, 1993; Schmitt et al., 2001; Schmitt, 2003). When there was no
gender difference in short-term mating likelihood, both women and men
had a likelihood of 25\%. When there was a gender difference in mating
standards, men had a mating standard of 3 (the highest possible value =
10), while women had a standard of 5 (Kenrick et al., 1990; Regan,
1998). When there was no gender difference in mating standards, both men
and women had a standard of 5.

@@

At each time step, men and women moved and decided on short-term mating
as the agents did in the heterosexual procedures. Half of the agents (75
men and 75 women) were also given the initiator status. The initiators
checked to see whether there were other agents at the same location. The
rest of the procedures were identical to the heterosexual procedures,
except that the agents only chose those with the same gender as
potential short-term mates or long-term partners.

@@

The model was run for 1,000 time steps in each simulation. We ran 10,000
simulations for each experiment, including 2,500 simulations for each
condition. We controlled for initializing random seeds in the model
runs. All simulations were run using Netlogo 6.2.1 (Wilensky, 1999).

@@

\hypertarget{gender-differences-in-short-term-mating-behaviors}{%
\subsection{Gender differences in short-term mating
behaviors}\label{gender-differences-in-short-term-mating-behaviors}}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=70mm]{figures/fig1_men_vs_women.pdf}
  \caption{\textbf{Short-term mating behaviors of men and women after 1,000 time steps in the model when gender differences existed in mating preferences.} Violin plots summarizing the outcome variables separately for heterosexual individuals and gay men and lesbian women. Plot (A) shows the average number of short-term mating experiences, and plot (B) shows the average number of short-term mates. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar. Statistics were calculated using the full population of men and women in the model (\textit{N\textsubscript{m}} = \textit{N\textsubscript{w}} = 150).}
\end{figure}

@@

In Experiment 1, however, when the means of the outcome variables were
calculated among men and women in the mating pool (the number of
individuals in the mating pool: \emph{M\textsubscript{m}} = 56.85,
\emph{SD\textsubscript{m}} = 5.99; \emph{M\textsubscript{w}} = 77.04,
\emph{SD\textsubscript{w}} = 7.53), there were gender differences in
short-term mating behaviors when men and women had differential
preferences for short-term mating. The average number of short-term
mating experiences was higher among heterosexual men (\emph{M} = 3.78,
\emph{SD} = 0.50) than among heterosexual women (\emph{M} = 2.79,
\emph{SD} = 0.35), \emph{t}(4,998) = -81.53, \emph{p} \textless{} .001,
Cohen's \emph{d} = 2.31. Likewise, the average number of short-term
mates was higher among heterosexual men (\emph{M} = 2.12, \emph{SD} =
0.17) than among heterosexual women (\emph{M} = 1.56, \emph{SD} = 0.11),
\emph{t}(4,998) = -135.85, \emph{p} \textless{} .001, Cohen's \emph{d} =
3.84.

\hypertarget{comparing-heterosexual-and-gay-men}{%
\subsection{Comparing heterosexual and gay
men}\label{comparing-heterosexual-and-gay-men}}

@@

In an exploratory manner, we also compared across conditions to see
which of the two dimensions of mating preferences contributed to gender
differences in short-term mating behaviors.

Among heterosexual individuals, gender differences in short-term mating
behaviors, as calculated among individuals in the mating pool, emerged
when men and women had different standards for short-term mates.
Heterosexual men (vs.~heterosexual women) in the mating pool had a
higher average number of short-term mating experiences and short-term
mates as long as they had lower mating standards, even when no gender
difference existed in short-term mating likelihood (experiences:
\emph{M\textsubscript{m}} = 2.82, \emph{SD\textsubscript{m}} = 0.36,
\emph{M\textsubscript{w}} = 2.21, \emph{SD\textsubscript{w}} = 0.28,
Cohen's \emph{d} = 1.89; partners: \emph{M\textsubscript{m}} = 1.76,
\emph{SD\textsubscript{m}} = 0.15; \emph{M\textsubscript{w}} = 1.38,
\emph{SD\textsubscript{w}} = 0.09, Cohen's \emph{d} = 3.13). In
comparison, virtually no gender differences in short-term mating
behaviors existed when men and women had the same mating standards
(Table 1).

@@

Among gay men and lesbian women, gender differences in short-term mating
behaviors emerged either when they had different short-term mating
likelihood or when their mating standards were different. Gay men
(vs.~lesbian women) had a higher average number of short-term mating
experiences and short-term mates when they had a higher short-term
mating likelihood (experiences: \emph{M\textsubscript{m}} = 0.80,
\emph{SD\textsubscript{m}} = 0.23, \emph{M\textsubscript{w}} = 0.32,
\emph{SD\textsubscript{w}} = 0.10, Cohen's \emph{d} = 2.67; partners:
\emph{M\textsubscript{m}} = 0.52, \emph{SD\textsubscript{m}} = 0.13,
\emph{M\textsubscript{w}} = 0.25, \emph{SD\textsubscript{w}} = 0.07,
Cohen's \emph{d} = 2.60), or when they had lower mating standards
(experiences: \emph{M\textsubscript{m}} = 1.05,
\emph{SD\textsubscript{m}} = 0.18, \emph{M\textsubscript{w}} = 0.32,
\emph{SD\textsubscript{w}} = 0.10, Cohen's \emph{d} = 5.10; partners:
\emph{M\textsubscript{m}} = 0.82, \emph{SD\textsubscript{m}} = 0.12,
\emph{M\textsubscript{w}} = 0.25, \emph{SD\textsubscript{w}} = 0.07,
Cohen's \emph{d} = 5.78). In comparison, virtually no gender differences
in short-term mating behaviors existed when gay men and lesbian women
had the same short-term mating likelihood and mating standards (Table
2).

@@

The present study aimed at using agent-based modeling to investigate
whether men and women's different preferences for short-term mating
resulted in any gender differences in short-term mating behaviors,
specifically, in the number of short-term mating experiences and
short-term mates, and if they did, under what circumstances. We raised
two hypotheses: 1) gay men would engage in more short-term mating
behaviors as compared to lesbian women, and 2) gay men would engage in
more short-term mating behaviors as compared to heterosexual men. The
results from 1,000 time steps in a model simulating men and women's
mating behaviors supported our hypotheses. First of all, as compared to
lesbian women, gay men had higher average numbers of short-term mating
experiences and short-term mates. Secondly, gay men also had higher
average numbers of short-term mating experiences and short-term mates as
compared to heterosexual men. In contrast, we found no gender
differences in short-term mating behaviors between heterosexual men and
women, although heterosexual men in the mating pool did have higher
average numbers of short-term mating experiences and short-term mates.

As we expected, heterosexual men and women did not differ in short-term
mating behaviors despite their differential preferences for short-term
mating. This was because heterosexual men and women had an equal total
number of short-term mating experiences , so as the total number of
short-term mates. Since the sex ratio was 1:1 in our model, the average
number of short-term mating experiences and short-term mates must be
equal between men and women as well. However, we did find that among
individuals in the mating pool, men engaged in more short-term mating
behaviors as compared to women. This was because there were more women
than men in the mating pool, resulting in lower averages among
heterosexual women despite the equal numbers of experiences and mates in
total.

Moreover, gender differences in short-term mating behaviors emerged when
heterosexual men and women had different mating standards, but not when
they had different short-term mating likelihood. When women had a higher
standard than men did, less men than women in the population were above
a potential partner's standard and thus had a chance to have sex with
them. This contributed to the unequal number of men and women in the
mating pool, which led to gender differences in short-term mating
behaviors among this population.

In the light of these results, the empirical observation of gender
differences in short-term mating behaviors among heterosexual
individuals (e.g., Petersen \& Hyde, 2010) appear to be perplexing
because this is illogical (e.g., Gurman, 1989). Possible explanations
for these empirical results have to highlight caveats in the observation
process. One possibility is that there was sampling bias in the
observations (e.g., Wiederman \& Dubois, 1998) since surveys regarding
short-term mating behaviors may tend to attract individuals who already
engage in such behaviors. Another possibility is that heterosexual men
and women tend to report short-term mating behaviors differently. This
difference may be a result of social desirability bias (Alexander \&
Fisher, 2003) or due to men and women's different estimation strategies
(e.g., men tend to round up) (Brown \& Sinclair, 1999).

Among gay men and lesbian women, large gender differences in short-term
mating behaviors existed when men and women had differential preferences
for short-term mating, which was in line with empirical observations
(e.g., Peplau et al., 1997, 2004). This was probably because the number
of short-term mating experiences and short-term mates no longer counted
towards men and women simultaneously. Any gender differences in mating
preferences would result in differences in behaviors. A closer look at
the results did support this postulation. Either a difference in the
likelihood of short-term mating or a difference in mating standards
alone could contribute to gender differences in mating behaviors. This
was perhaps because the former increased the probability of both parties
of a given gay couple deciding to have short-term mating, and the latter
increased the probability of both of them meeting each other's
standards, as compared to the case of a given lesbian couple.

Supporting the idea that gay men's preferences for short-term mating
were not constrained by those of women, we found that gay men engaged in
more short-term mating behaviors as compared to heterosexual men. This
was consistent with the empirical literature (e.g., Peplau et al.,
1997). Interestingly, gay men and heterosexual men had the same mating
likelihood and the same standards for short-term mates in our model. The
only difference was a change in the preferences of potential partners.
When men's partners had a stronger preferences for short-term mating
(i.e., having men vs.~having women as potential partners), men also
appeared to engage in more short-term mating behaviors.
